---
title: "Experiment 1 Replication with 4 New Items"
author: "Polina Tsvilodub"
date: "7/29/2019"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
# Experiment outline 
This script analyses the data from a comparison class elicitation pilot experiment with different adjective and target pais (n=80). 

We expect _more subordinate labels to occur in the incongruent condition_. In contrast, we expect _a similar proportion of superordinate and subordinate labels as in the previous experiment in the congruent condition_. Overall proportion of _subordinate labels should be higher in the predicative condition_. 
The proportion of superordinate labels is expected to be higher due to superordinate contextual competitors.

### Procedure
The pairings of the target and the adjective are altered in this experiment: i. e. in the congruent condition, if target is a chihuahua, the sentence includes the adjective "small", if it is a great dane, it includes "big". In the incongruent (critical) condition, if the target is a chihuahua, the sentence includes the adjective "big", if the target is a great dane, it includes "small" (between-subject).


## Data analysis
```{r}
library(tidyverse)
library(tidyboot)

# read in data
d <- read_csv('./../data/results_7_exp1-replica_10.csv')

#first look at the data
glimpse(d)

# comments
d_comments <- d %>%
  distinct(submission_id, problems, comments, fairprice)
d %>% group_by(condition, size) %>% count()

```

The participants recruited via MTurk were paid $1.10. 


##Spliting data into main and warm-up, excluding participants

6 participants are excluded because of their native language. 
There are participants using upper case in the bot captcha, but they are not excluded in this analysis. 
```{r}
# make sure how participants indicate their native language 
# sometimes participants use only "en" or "eng" for english
# excluded non-native speakers 
d %>% distinct(d$languages) %>% View()


d_filt <- d %>% 
  filter(grepl("English", languages, ignore.case = T)) %>%
  select(submission_id, trial_name, trial_number, size, item, botresponse, response, condition,  picture)

glimpse(d_filt)

d_filt %>% distinct(d_filt$botresponse) %>% View()
# extract main trials 
d_main <- d_filt %>% filter((trial_name =="main1")|(trial_name=="main2")) %>% 
  select(submission_id, trial_number, response,  size, item, condition, picture)

```

## Categorizing the data 
About 30 responses excluded due to task misundersatnding.
```{r}
# question1 and question2 are the sentence parts coded in the experiment 
# look at the different responses provided and categorize them 
d_main %>% distinct(d_main$response) %>% View()

d_main_valid <- subset(d_main, !(response %in% c("small"))) 

d_main_responseCat <- d_main_valid %>%
  rowwise() %>%
  mutate( # categorize responses 
    response_cat =
      ifelse(
      tolower(response) %in% c("bird", "dog", "fish", "flower",  "tree"), "super", "sub"),
    
    response_num = ifelse(response_cat == "sub", 1, 0),
    response_label = "sub"
  )
```
## Response category distribution
```{r}
d_main_responseCat %>% mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))) %>% ggplot(aes(x=response_cat)) +geom_bar(position=position_dodge()) + facet_wrap(~condition)
```

## Proportion of subordinate responses by condition, separated by congruency  

The proportion of subordinate responses is the dependent variable we test by manipulating the syntactic condition: "That's a big X" (prenominal) versus "That X is big" (predicative). 

Count of different conditions: about 160 incongruent observations, about 240 congruent observations. Given the task description, the confusion among participants seems to happens especially in incongruent trials.

```{r}
d_main_summary <- d_main_responseCat %>%
  group_by(condition, response_label) %>%
  tidyboot_mean(column = response_num) %>% # calculate proportion of subordinate labels in the different conditions 
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))
        )

ggplot(d_main_summary, aes(x = condition, fill = condition,
                           y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col(position = position_dodge(0.8))+
  geom_linerange(position = position_dodge(0.8))+
  labs( y = "Proportion subordinate responses")+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  ggtitle("The proportion of subordinate responses by syntactic condition")
 # facet_wrap(~condition)

```
## Stats
```{r}


library(lme4)
lm.rs <- glmer(response_num ~ condition + (1 | submission_id) + 
        (1 + condition | item), data = d_main_responseCat, family = 'binomial')
summary(lm.rs)
```

```{r}
cond_numeric <- ifelse(d_main_responseCat$condition=="prenominal", 1, -1)
lm.rsnum <- glmer(response_num ~ cond_numeric + (1 | submission_id) + 
        (1 + cond_numeric | item), data = d_main_responseCat, family = 'binomial')
summary(lm.rsnum)

```


## Proportion of subordinate labels in big vs. small trials 

We check if there is any proportional difference between the trials with different target size. The size does not seem to have a significant effect.  

```{r}
d_main_responseCat %>% group_by(size, response_label) %>% count()

d_main_summary_bySize <- d_main_responseCat %>%
  group_by(condition, size, response_label) %>%
  tidyboot_mean(column = response_num) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big")))

# big and small indicate the size of the target here 
ggplot(d_main_summary_bySize, aes(x = size, fill = size, 
                           y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col(position = position_dodge(0.8), width = 0.8)+
  geom_linerange(position = position_dodge(0.8))+
  labs( y = "Proportion subordinate responses")+
  facet_wrap(~condition)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+ ggtitle("Proportions in trials grouped by target size")
```

## Consistency of choosing a response category by subject

Do participants switch between superordinate and subordinate labels within the experiment?  


```{r}
d_main_responseCat %>%
  group_by(submission_id, condition, response_label) %>%
  summarize(n_sub_responses = sum(response_num)) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X (prenominal)", "That X is big (predicative)"))) %>%
  ggplot(., aes( x = n_sub_responses))+
  geom_bar(position=position_dodge())+
  facet_wrap(~condition) + ggtitle("Number of subordinate responses uttered per participant in the 6 trials")
```


## Proportion of subordinate responses by item (context)
For each context, there is a pair of targets (a big and a small one). The targets seem to elicit different proportions of subordinate labels. 
```{r}
d_main_responseCat %>% count(picture, condition) %>% View()



d_main_responseCat %>%
  group_by(condition, item, size, response_label) %>%
  tidyboot_mean(column = response_num) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))
         ) %>%

# big and small indicate the adjective actually used 
ggplot(., aes(x = size, fill = condition, 
                           y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col(position = position_dodge(0.8), width = 0.8)+
  geom_linerange(position = position_dodge(0.8))+
  labs( y = "Proportion subordinate responses")+
  facet_grid(~item)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))
# leave this grouping or interchange adj_cond and condition?

```

``` {r}
d_main_responseCat %>%
  group_by(condition, item,  response_label) %>%
  summarize(n_sub_responses = sum(response_num)) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))
         ) %>%

# big and small indicate the adjective actually used 
ggplot(., aes(x = n_sub_responses
                        ))+
  geom_bar(position = position_dodge(0.8), width = 0.8)+
 
  labs( y = "Proportion subordinate responses")+
  facet_grid(condition~item)
# leave this grouping or interchange adj_cond and condition?

```

``` {r}
d_main_responseCat %>%
  group_by(condition, item, size_target, response_label) %>%
  summarize(n_sub_responses = sum(response_num)) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))
         ) %>%

# big and small indicate the adjective actually used 
ggplot(., aes(x = n_sub_responses, fill = size_target
                        ))+
  geom_bar(position = position_dodge(0.8), width = 0.8)+
 
  labs( y = "Proportion subordinate responses")+
  facet_grid(condition~item)


```

``` {r}
d_main_responseCat.itemCon <- d_main_responseCat %>%
  group_by(picture, response_label) %>%
  summarize(n_sub_responses = sum(response_num),
            n = n(),
            prop = n_sub_responses / n) %>%
  ungroup() 


d_main_responseCat.itemCon %>%
  ggplot(., aes( x = prop ))+
  geom_histogram()+
  xlab("proportion subordinate responses")+
  ylab('number of items')
```

```{r}
d_main_responseCat.itemCon[with(d_main_responseCat.itemCon, order(prop)),] %>%
  select(picture, prop, n_sub_responses, n) %>%
  kable()

```