---
title: "Comparison Class Elicitation With Superordinate Context Items"
author: "Polina Tsvilodub"
date: "7/29/2019"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
#Experiment outline 
This script analyses the data from a comparison class elicitation pilot experiment with different adjective and target pais (n=80). 

We expect _more subordinate labels to occur in the incongruent condition_. In contrast, we expect _a similar proportion of superordinate and subordinate labels as in the previous experiment in the congruent condition_. Overall proportion of _subordinate labels should be higher in the predicative condition_. 

### Procedure
The pairings of the target and the adjective are altered in this experiment: i. e. in the congruent condition, if target is a chihuahua, the sentence includes the adjective "small", if it is a great dane, it includes "big". In the incongruent (critical) condition, if the target is a chihuahua, the sentence includes the adjective "big", if the target is a great dane, it includes "small".Every subject view three congruent and three incongruent conditions, randomly matched to three big and three small targets. 



## Data analysis
```{r}
library(tidyverse)
library(tidyboot)

# read in data
d <- read_csv('./../data/results_6_super-comp-class-exp_80.csv')

#first look at the data
glimpse(d)

# comments
d_comments <- d %>%
  distinct(submission_id, problems, comments, fairprice)

```

The participants recruited via MTurk were paid $0.50. 


##Spliting data into main and warm-up, excluding participants

6 participants are excluded because of their native language. 
There are participants using upper case in the bot captcha, but they are not excluded in this analysis. 
```{r}
# make sure how participants indicate their native language 
# sometimes participants use only "en" or "eng" for english
# excluded non-native speakers 
d %>% distinct(d$languages) %>% View()
d_langs <- d %>% mutate(
  languages = ifelse(languages == "yes", "English", languages)
) %>% mutate(
  languages = ifelse(is.na(languages), "X", languages)
) 

d_filt <- d %>% 
  filter(grepl("English", languages, ignore.case = T)) %>%
  select(submission_id, trial_name, trial_number, size_adj, item, botresponse, response, condition, size_target, adj_cond)

glimpse(d_filt)

d_filt %>% distinct(d_filt$botresponse) %>% View()
# extract main trials 
d_main <- d_filt %>% 
  select(submission_id, trial_number, response, size_adj, item, condition, size_target, adj_cond)

```

## Categorizing the data 
Two responses to be excluded from the analysis: "herd of primates", "collection of plants"
```{r}
# question1 and question2 are the sentence parts coded in the experiment 
# look at the different responses provided and categorize them 
d_main %>% distinct(d_main$response) %>% View()

d_main_valid <- subset(d_main, !(response %in% c("horse", "house", "koala", "butterfly", "squirrel", "shark", "whale", "flamingo", "snake", "rose", "crab", "pug", "tiger", "tail", "parade", "fox", "deer", "item"))) 
d_main_valid$response[d_main_valid$response == "gold fish"] <- "goldfish"
d_main_valid$response[d_main_valid$response == "straight tail monkey"] <- "monkey"
d_main_valid$response[d_main_valid$response == "scaly fish"] <- "fish"
d_main_valid$response[d_main_valid$response == "blue and white fish"] <- "fish"
d_main_valid$response[d_main_valid$response == "grey dog"] <- "dog"

  
d_main_responseCat <- d_main_valid %>%
  rowwise() %>%
  mutate( # categorize responses 
    response_cat =
      ifelse(
      tolower(response) %in% c("ape", "primate", "bird", "dog", "fish", "flower", "monkey", "tree", "leaf", "twig", "branch", "bush"), "basic", "sub"),
    
    response_num = ifelse(response_cat == "sub", 1, 0),
    response_label = "sub"
  )
d_main_responseCat$response_cat[d_main_responseCat$response == "plant"] <- "super"
d_main_responseCat$response_cat[d_main_responseCat$response == "animal"] <- "super"
```
## Response category distribution
```{r}
d_main_responseCat %>% mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))) %>% ggplot(aes(x=response_cat, fill=adj_cond)) +geom_bar(position=position_dodge()) + facet_wrap(~condition)
```

## Proportion of subordinate responses by condition, separated by congruency  

The proportion of subordinate responses is the dependent variable we test by manipulating the syntactic condition: "That's a big X" (prenominal) versus "That X is big" (predicative). 

Count of different conditions: about 160 incongruent observations, about 240 congruent observations. Given the task description, the confusion among participants seems to happens especially in incongruent trials.

```{r}
d_main_summary <- d_main_responseCat %>%
  group_by(condition, response_label, adj_cond) %>%
  tidyboot_mean(column = response_num) %>% # calculate proportion of subordinate labels in the different conditions 
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big")),
         adj_cond = factor(adj_cond, 
                            levels=c("congruent", "incongruent"), 
                           labels = c("great dane : big", "great dane : small")))

ggplot(d_main_summary, aes(x = adj_cond, fill = adj_cond,
                           y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col(position = position_dodge(0.8))+
  geom_linerange(position = position_dodge(0.8))+
  labs( y = "Proportion subordinate responses")+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  ggtitle("The proportion of subordinate responses by syntactic condition")+
  facet_wrap(~condition)

```

## Proportion of subordinate responses by congruence 
```{r}
d_main_congr <- d_main_responseCat %>%
  group_by(adj_cond, condition, response_label) %>%
  tidyboot_mean(column = response_num) %>% # calculate proportion of subordinate labels in the different conditions 
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big")),
         adj_cond = factor(adj_cond, 
                            levels=c("congruent", "incongruent"), 
                           labels = c("great dane : big", "great dane : small")))

ggplot(d_main_congr, aes(x = condition, fill = condition,
                           y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col(position = position_dodge(0.8))+
  geom_linerange(position = position_dodge(0.8))+
  labs( y = "Proportion subordinate responses")+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  ggtitle("Proportion of subordinate responses by congruence")+
  facet_wrap(~adj_cond)
```

## Proportion of subordinate labels in big vs. small trials 

We check if there is any proportional difference between the trials with different target size. The size does not seem to have a significant effect.  

```{r}
d_main_responseCat %>% group_by(size_target, adj_cond) %>% count()

d_main_summary_bySize <- d_main_responseCat %>%
  group_by(condition, size_target, adj_cond, response_label) %>%
  tidyboot_mean(column = response_num) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big")),
         adj_cond = factor(adj_cond, 
                            levels=c("congruent", "incongruent"), 
                           labels = c("great dane : big", "great dane : small")))

# big and small indicate the size of the target here 
ggplot(d_main_summary_bySize, aes(x = size_target, fill = adj_cond, 
                           y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col(position = position_dodge(0.8), width = 0.8)+
  geom_linerange(position = position_dodge(0.8))+
  labs( y = "Proportion subordinate responses")+
  facet_wrap(~condition)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+ ggtitle("Proportions in trials grouped by target size")
```
Interestingly, small targets seem to elicit less subordinates in the predicative condition (more sensistive to syntactic structure), whereas big targets elicit equal proportions in both conditions. Is there a different prior of using positive adjectives ('big' as the incongruent one in small trials) (in terms of informativity) than using negative adjectives (incongruent ones in the big trials)?  Is it 'worse' to say "big chihuahua" in comparison to dogs than "small great dane" in comparison to dogs?
Also, the proportions reverse in the congruent conditions depending on the target size! This should not be the case! People rather say "That's a big great dane" (>0.5), but perform as expected in the predicative condition (0.5). Again, different prior expectations of positive vs negative feature expression on a scale? 
In the small trials, they perform as expected (predicative > prenominal).


## Consistency of choosing a response category by subject

Do participants switch between superordinate and subordinate labels within the experiment?  

Add statistics on no/all responses of same cat, more/less than half/mixed
```{r}
d_main_responseCat %>%
  group_by(submission_id, adj_cond, condition, response_label) %>%
  summarize(n_sub_responses = sum(response_num)) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X (prenominal)", "That X is big (predicative)")),
         adj_cond = factor(adj_cond, 
                            levels=c("congruent", "incongruent"), 
                           labels = c("great dane : big", "great dane : small"))) %>%
  ggplot(., aes( x = n_sub_responses, fill = adj_cond))+
  geom_bar(position=position_dodge())+
  facet_grid(adj_cond~condition) + ggtitle("Number of subordinate responses uttered per participant in the 3 trials")
```
There are many people in the congruent condition consistently using subordinates in the prenominal condition, but also some not using them at all. Incongruent trials seem to be as expected.  

## Proportion of subordinate responses by item (context)
For each context, there is a pair of targets (a big and a small one). The targets seem to elicit different proportions of subordinate lables. The most effective ones seem to be the swan, the chihuahua and the bonsai. However, the results are not robust due to a small sample size. 
```{r}
d_main_responseCat %>% count(item, condition) %>% glimpse()



d_main_responseCat %>%
  group_by(condition, item, size_adj, size_target, response_label) %>%
  tidyboot_mean(column = response_num) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))
         ) %>%

# big and small indicate the adjective actually used 
ggplot(., aes(x = size_target, fill = size_adj, 
                           y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_col(position = position_dodge(0.8), width = 0.8)+
  geom_linerange(position = position_dodge(0.8))+
  labs( y = "Proportion subordinate responses")+
  facet_grid(condition~item)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))
# leave this grouping or interchange adj_cond and condition?

```
``` {r}
d_main_responseCat %>%
  group_by(condition, item,  response_label) %>%
  summarize(n_sub_responses = sum(response_num)) %>%
  ungroup() %>%
  mutate(condition = factor(condition, 
                            levels = c("prenominal", "predicative"),
                            labels= c("That's a big X", "That X is big"))
         ) %>%

# big and small indicate the adjective actually used 
ggplot(., aes(x = n_sub_responses, 
                        ))+
  geom_bar(position = position_dodge(0.8), width = 0.8)+
 
  labs( y = "Proportion subordinate responses")+
  facet_grid(condition~item)
# leave this grouping or interchange adj_cond and condition?

```
